A crossover code for high-dimensional composition

### 2019-09-13 ms_neurips_2019/ms_neurips_2019.tex
Do final revisions before submission.

% ready for submission
% \usepackage{neurips_2019}

@@

% to compile a camera-ready version, add the [final] option, e.g.:
     \usepackage[final]{neurips_2019}

@@

An ideal code should allow fast readout of both a composition's base elements and their relations, but most HDC schemes support one or the other. Sums of random HD vectors, for instance, have high overlap (low distance under most metrics) to their elements, enabling element readout via overlaps with the sum \cite{Bloom:1970, Anderson:1973, Plate:1994}; but cannot encode relations. HDC operations like circular convolution \cite{Plate:1995}, permutation \cite{Gayler:1998, Sahlgren:2008}, or matrix multiplication \cite{Gosmann:2019}, yield codes with cue-recoverable relations (cue A recovers B from ((A,B),(C,D))), but low overlap to the composition's base elements, precluding fast element readout without cues (but see \cite{Rachkovskij:2001}); this also conflicts with our desire that AB's code should resemble A's more than C's. Here we present a new HD composition operation inspired by chromosomal crossover, where two vectors $X$ and $X'$ are ordered or bound by masking out a fraction of $X$'s components and replacing them with those from $X'$; relational information is encoded in the context-dependent mask $\mu(X, X')$. "Crossover" codes thus highly overlap with their base elements, allowing fast recovery of both elements and relations, and decoding by greedy reconstruction. Further, they evenly distribute information and exhibit object-complexity-invariant statistics, ideal for robust intelligent systems. We describe crossover for sequences first, then generalize to binding and trees.

@@

Let $X$ be a vector of $N$ components, each of which can take any of $Z$ states, $x_j \in G \equiv \{1,...,Z\}$, and let $d(X, X')$ be the Hamming distance from $X$ to $X'$. For a symbol sequence $Y \equiv Y_{1:L} \equiv (y_1, ..., y_L)$, where $y_t \in D \equiv \{1, ..., M\}$, we establish a mapping from $Y$ to $X^Y$ as follows:

@@

Using $C$ and $\mu$ we encode sequences as in Algorithm 1 and Figure 1a: starting with $X^*$, we sequentially "weave" in symbol codes $X^{y_1}$, .., $X^{y_L}$ via an operation "CrossoverSeq": at each $t$ we mask out a fraction $1/(t+1)^\gamma$ of the evolving code's components and replace them with those from $X^{y_t}$. The mask selects components where $\mu_j(X^{Y_{1:t-1}}, X^{y_t}) < 1/(t+1)^\gamma$, so is unique for all $(X^{Y_{1:t-1}}, X^{y_t})$, yielding context-dependent mixing. $\gamma \geq 1$ prevents decay of early symbols in $X^Y$.

@@

\State $mask \gets (\mu(X, X') < 1/(t+1)^\gamma) \quad$  \# context-dependent mask with about $1/(t+1)^\gamma$ 1's
\State $X'' \gets X$
\State $X''[mask] \gets X'[mask] \quad$  \# replace $X$'s masked components with those from $X'$

@@

\begin{figure}
  \centering
  \includegraphics{1_sequences.pdf}
  \caption{Sequence encoding via crossover ($N=1024, Z=8, M=26, \gamma=1$). (a) Example encoding of CAFE; colors are states, inscribed symbols are visual aids, not used by the algorithm. (b) Sequence-symbol code distances. (c) State distributions for example short and long sequence for one $C, \mu$ sample. (d, e) Distances of CAFE's and CICADA's code to other sequence codes. (f) Decoding vs code corruption. "x"s indicate theoretical predictions and bars average simulation results. All plots here and elsewhere were made via 100 samples of $C, \mu$. "*" prefixes indicate sequences.}
  \label{fig:1}
\end{figure}

For large $Z$ and $\gamma = 1$, mean symbol-sequence similarity (over $C, \mu$) scales with symbol $i$'s count, $n^i$, within the sequence (Figure 1b), with variance $\sim 1/N$. Specifically, $d(X^Y, X^i) \sim \textrm{Binomial}(N, 1-q)/N$, with $q = P(x^Y_j = x^i_j) = n^i/(L+1) + (1/Z)(L+1-n^i)/(L+1)$. As exemplified by Figure 1c, crossover code components are uniform over $G$, i.i.d. for a given $Y$, and invariant to sequence statistics (e.g. length). The proof follows from the i.i.d. nature of $C, \mu$ and the interchangeability of component states, and suggests crossover as a robust, scale-invariant encoding.

For two sequences $Y$ and $Y'$ sharing their first $t$ symbols, $d(X^Y, X^{Y'})$ is also a scaled Binomial variable, and $q$ can be computed exactly for arbitrary $\gamma$; we present the $\gamma = 1$ case here. Let $\mathbf{n}_u$ be the length-$M$ symbol count vector for $Y_{1:t} = Y'_{1:t}$ with $\mathbf{1}^T\mathbf{n}_u = t$, and $\mathbf{n}_v$ and $\mathbf{n}_{v'}$ be the symbol count vectors for $Y_{t+1:L}$ and $Y'_{t+1:L'}$, respectively, with $\mathbf{1}^T\mathbf{n}_{v} = L - t$ and $\mathbf{1}^T\mathbf{n}_{v'} = L' - t$. Then:
$$q \equiv P(x_j^Y = x_j^{Y'}) = \frac{t+1}{L+1}\frac{t+1}{L'+1} + \frac{1}{L+1}\frac{1}{L'+1}\left[\mathbf{n}^T_v\mathbf{n}_{v'} + \left((L-t)(L'-t) - \mathbf{n}^T_v\mathbf{n}_{v'} \right)\frac{1}{Z}\right]$$
$$+\frac{1}{L+1}\frac{1}{L'+1}\left[
\mathbf{n}^T_{u}\mathbf{n}_{v'} + \left((t+1)(L'-t) - \mathbf{n}^T_{u}\mathbf{n}_{v'} \right)\frac{1}{Z}
+ \mathbf{n}^T_{v}\mathbf{n}_{u} + \left((L-t)(t+1) - \mathbf{n}^T_{v}\mathbf{n}_{u} \right)\frac{1}{Z}
\right]$$
For most $Y$, $X^Y$ overlaps little with codes for $Y$'s anagrams, and more with correct than incorrect starting subsequences, e.g. CI's and CIC's codes are closer than CD's and CIA's codes, respectively, to CICADA's code (Figure 1d,e). This is because different next-symbols yield different $\mu$, which quickly decreases overlap with $X^Y$ for sequence codes with out-of-order symbols. Consequently, $Y$ can "greedily" decoded by rebuilding $X^Y$ from $X^*$, outputting at each rebuild step $t$ the symbol $i$ maximizing overlap with $X^Y$ when $X^i$ is woven into the current code, then weaving in $X^i$ and advancing to $t+1$. Decoding is robust to noise added to $X^Y$, and when a prior $P(Y)$ exists, the mutual information between $x_j^Y$ and $Y$ is i.i.d. across $j$; robustness is thus equivalent to random and targeted attacks, so decoding accuracy depends only on the number of corrupted $x_j^Y$ (Figure 1f).

While crossover intrinsic supports sequences with repeated symbols (either consecutive or not), for $Y$ with many symbol repeats distinct from $y_1$ (e.g. REFEREE), $X^*$ crossed over with $X^{y_1}$ can be further from $X^Y$ than $X^*$ crossed with the repeated symbol, impairing decoding. Making $\gamma > 1$ fixes this by more heavily weighting early symbols (Figure 2a). Surprisingly, while this front-loaded weighting improves decoding of, say, ABBBBBB, it does not equivalently degrade decoding of AAAAAAB, and optimal $\gamma$ are shared by many sequences (Figure 2b). In general, decoding accuracy decreases (increases) with $L$ ($N$), since $\textrm{Var}[P(x_j^{Y_{1:t}} = x_j^Y)] \sim 1/N$. Decoding accuracy is also robust to large $M$, the dictionary size, due at heart to the low expected overlap of HD random vectors.

\begin{figure}
\label{fig:2}
  \centering
  \includegraphics{2_gamma.pdf}
  \caption{Crossover modifications ($N=1024, Z=8, M=26$). (a) Distance of *REFEREE to symbols/sequences for two $\gamma$. (b) Decoding vs. $\gamma$ for several sequences. (c) Decoding of *CAFE vs number of corrupted components at encoding step for i.i.d. and correlated $\mu$. (d) Decoding of *CICADA for i.i.d., correlated $\mu$, and correlated $\mu$ with increased $N$.}
\end{figure}

While crossover with i.i.d. $\mu$ is robust to noise added to the final code (Figure 1f), it fails when noise is added during encoding (Figure 2c,d), as one altered component yields an entirely different mask. This is fixed by introducing correlations into $\mu_j(X, X')$ (here by making $\mu$ a function of sparse components of $X, X'$) (Figure 2c,d). The cost of this is decreased decoding accuracy for certain sequences, since correlated masks are less distinct, but this can be corrected by increasing $N$.

Beyond sequences, crossover naturally generalizes to \textit{k}-ary trees. Key to encoding trees is \textit{binding} elements into groups that can themselves be bound, yet without forgetting initial group identities: ((A,B), (C,D)) must be encoded differently from ((A,D), (C,B)). In crossover, $k$ elements $i_0, ..., i_k$ can be bound into a group by using a context-dependent mask function $\mu_j(X^{i_0}, .... X^{i_k})$ (optionally commuting). The vector for the group takes components from $X^{i_0}$ where $0 < \mu < 1/k$, from $X^{i_1}$ where $1/k < \mu < 2/k$, etc. This can then be bound itself in the exact same manner (Figure 3a), allowing recursive encoding of an arbitrary tree. The result is a vector overlapping more with elements in vs not in the tree, and with groups in vs not in tree, even if the latter contain the same base elements (Figure 3b). A generic tree encoding will tend to overlap more with its included sub-trees than with alternative sub-trees of equal complexity. Similar to sequence encoding, this means both base elements and their recursive relations can be extracted via their overlap with the final crossover code, and that decoding of the full tree can occur through a recursive, greedy reconstruction.

@@

While intelligent systems must learn at least some environment statistics to operate, it can be useful for certain computational features to be built in \cite{Zador:2019}. Flexible compositional coding is of particular relevance as it provides a substrate for computing with novel objects or events if they are made from familiar elements: an autonomous vehicle should be able to construct and manipulate a representation of "child on scooter behind car in front of bus", even if previously it had only learned the isolated scene elements. While many trained systems, e.g. translation networks \cite{Bahdanau:2014, Luong:2015, Wu:2016}, have implicit capacity for compositional encoding (otherwise outputs would be disordered), it is unclear how they could handle complexity beyond their training data, and understanding their internal representations can be challenging \cite{Lipton:2016} (although exciting progress is being made in interpretability \cite{Zeiler:2014, Montavon:2018}). Alternatively, one can take the HDC approach of designing algorithms with \textit{a priori} flexibility, then attach or embed them in trainable systems. Basing training on convolutional binding, for instance, allowed an artificial network to efficiently and scalably learn knowledge graphs and out-perform state-of-the-art methods at link prediction \cite{Nickel:2016}. It stands to reason that identifying robust HDC codes could prove useful to improving artificial systems, as well as to guiding research in neural computation in biology.

"Crossover" shares advantages with existing HDC schemes but is distinct in key ways also. Like other schemes \cite{Kanerva:2009}, crossover preserves recoverable relational information, and relation codes can be reused as elements, allowing arbitrary deep composition. For instance, in crossover one can add CAFE's or other words' codes to the dictionary (perhaps with minor noise added), enabling encoding of long sequences made from short ones. To encode relations, previous schemes have used circular convolution \cite{Plate:1995}, XOR (for binary vectors) \cite{Kanerva:1994}, permutation \cite{Sahlgren:2008, Gayler:1998}, or matrix multiplication \cite{Gosmann:2019}, which yield relation codes that little overlap their elements. While these distinguish different compositions of the same elements and allow cue-based recall, base elements cannot be read out directly from their overlap with the relation code, and the entire dictionary must be used (instead of just the relevant symbols) when reading out relations, potentially greatly slowing decoding for large $M$. An exception is "context-dependent thinning" (CDT), which stores relations via the union of the elements, then "stamps" the code with a context-dependent set of zeros \cite{Rachkovskij:2001}; while the stamp shape retains relational information without erasing element overlaps, however, information beneath its imprint is discarded, and it is not clear how CDT could construct recursive sequence codes without early elements decaying exponentially. Crossover overcomes these limitations, as both elements and their relations can be read out directly, and in sequences no element overlaps decay exponentially. Further, crossover yields i.i.d. information across code components and code statistics invariant to object complexity, suggesting it may be a useful representation for a robust intelligent system.

Crossover is also conceptually distinct from its predecessors since one need not define arithmetical operations over code components. Whereas previous HDC algorithms require summing components, either directly or through convolution \cite{Plate:1995, Gayler:1998, Sahlgren:2008, Gosmann:2019, Rachkovskij:2001, Kanerva:1994}, crossover's component states $1, ..., Z$ are just labels, with no numeric value; they need only a comparison operation. Crossover thus outlines a distributed computation substrate that can exist outside vector spaces, and instead within more generic metric spaces. This is relevant to biological implementation, as it allows one to envision $x_j$ not only as firing rates of neurons or populations, but also as neural assembly states that might not make sense to add, multiply, or rank. For instance, each component of a crossover code could correspond to the attractor state of a small Hopfield-like assembly \cite{Hopfield:1982}, with the collective state of a pool of $N$ such assemblies encoding a sequence or tree. Such could serve, for instance, as a crude implementation of a combinatorially expressive working memory substrate, yet with sufficient code overlap statistics to be usefully read out and manipulated by downstream and recurrent computations.

\subsubsection*{Acknowledgments}

Use unnumbered third level headings for the acknowledgments. All acknowledgments
go at the end of the paper. Do not include acknowledgments in the anonymized
submission, only in the final paper.

>>>>

% ready for submission
\usepackage{neurips_2019}

@@

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2019}

@@

An ideal code should allow fast readout of both a composition's base elements and their relations, but most HDC schemes support one or the other. Sums of random HD vectors, for instance, have high overlap (low distance under most metrics) to their elements, enabling element readout via overlaps with the sum \cite{Bloom:1970, Anderson:1973, Plate:1994}; but cannot encode relations. HDC operations like circular convolution \cite{Plate:1995}, permutation \cite{Gayler:1998, Sahlgren:2008}, or matrix multiplication \cite{Gosmann:2019}, yield codes with cue-recoverable relations (cue A recovers B from ((A,B),(C,D))), but low overlap to the composition's base elements, precluding fast element readout without cues (but see \cite{Rachkovskij:2001}); this also conflicts with our desire that AB's code should resemble A's more than C's. Here we present a new HD composition operation inspired by chromosomal crossover, where two vectors $X$ and $X'$ are ordered or bound by masking out a fraction of $X$'s components and replacing them with those from $X'$; relational information is encoded in the context-dependent mask $\mu_:(X, X')$. "Crossover" codes thus highly overlap with their base elements, allowing fast recovery of both elements and relations, and decoding by greedy reconstruction. Further, they evenly distribute information and exhibit object-complexity-invariant statistics, ideal for robust intelligent systems. We describe crossover for sequences first, then generalize to binding and trees.

@@

Let $X$ be a vector of $N$ components, each of which can take any of $Z$ states: $x_j \in G \equiv \{1,...,Z\}$; and let $d(X, X')$ be the Hamming distance from $X$ to $X'$. For a symbol sequence $Y \equiv Y_{1:L} \equiv (y_1, ..., y_L)$, where $y_t \in D \equiv \{1, ..., M\}$, a dictionary of $M$ symbols, we map $Y$ to $X^Y$ as follows:

@@

Using $C$ and $\mu$ we encode sequences as shown in Algorithm 1 and Figure 1a: starting with $X^*$, we sequentially "weave" in symbol codes $X^{y_1}$, .., $X^{y_L}$ via an operation "CrossoverSeq": at each $t$ we mask out a fraction $1/(t+1)^\gamma$ of the evolving code's components and replace them with those from $X^{y_t}$. The mask selects components where $\mu_j(X^{Y_{1:t-1}}, X^{y_t}) < 1/(t+1)^\gamma$, so is unique for all $(X^{Y_{1:t-1}}, X^{y_t})$, yielding context-dependent mixing. $\gamma \geq 1$ prevents decay of early symbols in $X^Y$.

@@

\State $mask \gets (\mu_:(X, X') < 1/(t+1)^\gamma) \quad$  \# context-dependent mask with about $1/(t+1)^\gamma$ 1's
\State $X'' \gets X$
\State $X''[mask] \gets X'[mask] \quad$  \# replace $X$'s masked components with those from $X'$

@@

\begin{figure}
  \centering
  \includegraphics{1_sequences.pdf}
  \caption{Sequence encoding via crossover ($N=1024, Z=8, M=26, \gamma=1$). (a) Example encoding of CAFE; colors are states; inscribed symbols are visual aids, not used by the algorithm. (b) Sequence-symbol code distances. (c) State distributions for example short and long sequence for one $C, \mu$ sample. (d, e) Distances of CAFE's and CICADA's code to other sequence codes. (f) Decoding vs code corruption. All plots here/elsewhere were made via 100 $C, \mu$ samples. "x"s denote exact calculations and bars average simulation results. "*" prefixes distinguish sequences from symbols.}
  \label{fig:1}
\end{figure}

For large $Z$ and $\gamma = 1$, mean similarity ($1-d$) of a symbol $i$ to the final sequence scales with symbol $i$'s count, $n^i$, within the sequence (Figure 1b), with variance $\sim 1/N$. Specifically, $d(X^Y, X^i) \sim \textrm{Binomial}(N, 1-q)/N$, with $q = P(x^Y_j = x^i_j) = n^i/(L+1) + (1/Z)(L+1-n^i)/(L+1)$. As exemplified by Figure 1c, crossover code components are uniform over $G$, i.i.d. for a given $Y$, and invariant to sequence statistics (e.g. length). The proof follows from the i.i.d. nature of $C, \mu$ and the interchangeability of component states, and suggests crossover as a robust, scale-invariant encoding.

For two sequences $Y$ and $Y'$ sharing their first $t$ symbols, $d(X^Y, X^{Y'})$ is also a scaled Binomial variable, and $q$ can be computed exactly for arbitrary $\gamma$; we present the $\gamma = 1$ case here. Let $\mathbf{n}_u$ be the length-$M$ symbol count vector for $Y_{1:t} = Y'_{1:t}$ with $\mathbf{1}^T\mathbf{n}_u = t$, and $\mathbf{n}_v$ and $\mathbf{n}_{v'}$ be the symbol count vectors for $Y_{t+1:L}$ and $Y'_{t+1:L'}$, respectively, with $\mathbf{1}^T\mathbf{n}_{v} = L - t$ and $\mathbf{1}^T\mathbf{n}_{v'} = L' - t$. Then:
$$q \equiv P(x_j^Y = x_j^{Y'}) = \frac{t+1}{L+1}\frac{t+1}{L'+1} + \frac{1}{L+1}\frac{1}{L'+1}\left[\mathbf{n}^T_v\mathbf{n}_{v'} + \left((L-t)(L'-t) - \mathbf{n}^T_v\mathbf{n}_{v'} \right)\frac{1}{Z}\right]$$
$$+\frac{1}{L+1}\frac{1}{L'+1}\left[
\mathbf{n}^T_{u}(\mathbf{n}_{v} + \mathbf{n}_{v'}) + \left((t+1)(L'+L-2t) - \mathbf{n}^T_{u}(\mathbf{n}_{v} + \mathbf{n}_{v'})
 \right)\frac{1}{Z}
\right]$$
For most $Y$, $X^Y$ overlaps little with codes for $Y$'s anagrams, and more with correct than incorrect starting subsequences, e.g. CI's and CIC's codes are closer than CD's and CIA's codes, respectively, to CICADA's code (Figure 1d,e). This is because different next-symbols yield different $\mu$, which quickly decreases overlap with $X^Y$ for sequence codes with out-of-order symbols. Consequently, $Y$ can be "greedily" decoded by rebuilding $X^Y$ from $X^*$, outputting at each rebuild step $t$ the symbol $i$ that maximizes overlap with $X^Y$ when $X^i$ is woven into the current code, then weaving in $X^i$ and advancing to $t+1$. Decoding is robust to noise added to $X^Y$, and when a prior $P(Y)$ exists, the mutual information between $x_j^Y$ and $Y$ is i.i.d. across $j$; robustness is thus equivalent for random and targeted attacks, with decoding accuracy depending only on the number of corrupted $x_j^Y$ (Figure 1f).

While crossover intrinsically supports sequences with repeated symbols (either consecutive or not), for $Y$ with many symbol repeats distinct from $y_1$ (e.g. REFEREE), $X^*$ crossed over with $X^{y_1}$ can be further from $X^Y$ than $X^*$ crossed with the repeated symbol, impairing decoding. Making $\gamma > 1$ fixes this by more heavily weighting early symbols (Figure 2a). Surprisingly, while this front-loaded weighting improves decoding of, say, ABBBBBB, it does not equivalently degrade decoding of AAAAAAB, and optimal $\gamma$ are shared by many sequences (Figure 2b). In general, decoding accuracy decreases (increases) with $L$ ($N$), since $\textrm{Var}[P(x_j^{Y_{1:t}} = x_j^Y)] \sim 1/N$. Decoding accuracy is also robust to large $M$, the dictionary size, due at heart to the low expected overlap of HD random vectors.

\begin{figure}
\label{fig:2}
  \centering
  \includegraphics{2_gamma.pdf}
  \caption{Crossover modifications ($N=1024, Z=8, M=26$). (a) Distance of *REFEREE to symbols/sequences for two $\gamma$. (b) Decoding vs. $\gamma$ for several sequences. (c) Decoding of *CAFE vs number of corrupted components at encoding step for i.i.d. and correlated $\mu$. (d) Decoding of *CICADA for i.i.d., correlated $\mu$, and correlated $\mu$ with quadrupled $N$.}
\end{figure}

While crossover with i.i.d. $\mu$ is robust to noise added to the final code (Figure 1f), it fails when noise is added during encoding (Figure 2c,d), as one altered component yields an entirely different mask. This is fixed by introducing correlations into $\mu_j(X, X')$ (i.e. making $\mu$ a smooth function of $X, X'$ [many methods work]). The cost of this is decreased decoding accuracy for certain sequences, since correlated are less distinct than i.i.d. masks, but this can be corrected by increasing $N$ (Figure 2c,d).

Beyond sequences, crossover naturally generalizes to \textit{k}-ary trees. Key to encoding trees is \textit{binding} elements into groups that can themselves be bound, yet without forgetting initial group identities: ((A,B), (C,D)) must be encoded differently from ((A,D), (C,B)). In crossover, $k$ elements $i_1, ..., i_k$ can be bound into a group using a context-dependent mask function $\mu_j(X^{i_1}, .... X^{i_k})$ (which can optionally commute). The vector for the group takes components from $X^{i_1}$ where $0 < \mu < 1/k$, from $X^{i_2}$ where $1/k < \mu < 2/k$, etc. This can then be bound itself in the exact same manner (Figure 3a), allowing recursive encoding of an arbitrary tree. The result is a vector overlapping more with elements in vs not in the tree, and with groups in vs not in tree, even if the latter contain the same base elements (Figure 3b). A generic tree encoding will also tend to overlap more with its included sub-trees than with alternative sub-trees of equal complexity. Similar to sequence encoding, this means both base elements and their recursive relations can be extracted via their overlap with the final code, and that decoding of the full tree can occur through recursive but greedy reconstruction.

@@

While intelligent systems must learn at least some environment statistics to operate, it can be useful for certain computational features to be built in \cite{Zador:2019}. Flexible compositional coding is of particular relevance as it provides a substrate for computing with novel objects or events if they are made from familiar elements: an autonomous vehicle should be able to manipulate a representation of "child on scooter behind car in front of bus", even if previously it had only worked with the isolated scene elements. While many trained systems, e.g. translation networks \cite{Bahdanau:2014, Luong:2015, Wu:2016}, have some implicit capacity for compositional encoding (otherwise outputs would be disordered), it is unclear how they could handle complexity beyond their training data, and understanding their internal representations can be challenging \cite{Lipton:2016} (although exciting progress is being made in interpretability \cite{Zeiler:2014, Montavon:2018}). Alternatively, one can take the HDC approach of designing algorithms with \textit{a priori} flexibility, then attach or embed them in trainable systems. Basing training on convolutional binding, for instance, allowed an artificial network to efficiently and scalably learn knowledge graphs and out-perform state-of-the-art methods at link prediction \cite{Nickel:2016}. It stands to reason that identifying robust HDC codes could prove useful to improving artificial systems, as well as to guiding investigation of neural computation in biology.

"Crossover" shares advantages with existing HDC schemes but is distinct in key ways also. As in other schemes \cite{Kanerva:2009}, crossover preserves recoverable relational information, and relation codes can be reused as elements, allowing arbitrary deep composition. For instance, in crossover one can add CAFE's or other words' codes to the dictionary (perhaps with minor noise added), enabling encoding of long sequences made from short ones. To encode relations, previous schemes have used circular convolution \cite{Plate:1995}, XOR (for binary vectors) \cite{Kanerva:1994}, permutation \cite{Sahlgren:2008, Gayler:1998}, or matrix multiplication \cite{Gosmann:2019}, which yield relation codes that little overlap their elements. While these distinguish different compositions of the same elements and allow cue-based recall, base elements cannot be read out directly from their overlap with the relation code, and the entire dictionary must be used (instead of just the relevant symbols) when reading out relations, potentially greatly slowing decoding for large $M$. An exception is "context-dependent thinning" (CDT), which stores relations via the union of the elements, then "stamps" the code with a context-dependent set of zeros \cite{Rachkovskij:2001}; while the stamp shape retains relational information without erasing element overlaps, however, information beneath its imprint is discarded, and it is not clear how CDT could construct recursive sequence codes without early elements decaying exponentially. Crossover overcomes these limitations, as both elements and their relations can be read out directly, and in sequences no element overlaps decay exponentially. Further, crossover yields i.i.d. information across code components, and code statistics are invariant to object complexity, suggesting it may be a useful representation for a robust intelligent system.

Crossover is also conceptually distinct from its predecessors since one need not define arithmetical operations over code components. Whereas previous HDC algorithms require \textit{summing} components, either directly or through convolution \cite{Plate:1995, Gayler:1998, Sahlgren:2008, Gosmann:2019, Rachkovskij:2001, Kanerva:1994}, crossover's component states $1, ..., Z$ are just labels, with no numeric value; they need only a comparison operation. Crossover thus outlines a distributed computation substrate that can exist outside vector spaces, and instead within more generic metric spaces. This is relevant to biological implementation, as it allows one to envision $x_j$ not only as firing rates of neurons or populations, but also as neural assembly states that might not make sense to add, multiply, or rank. For instance, each component of a crossover code could correspond to the attractor state of a small Hopfield-like assembly \cite{Hopfield:1982}, with the collective state of a pool of $N$ such assemblies encoding a sequence or tree. This could serve, for instance, as a crude implementation of a combinatorially expressive working memory substrate, yet with sufficient code-overlap statistics to be usefully read out and manipulated by further downstream and recurrent computations.

% \subsubsection*{Acknowledgments}

% Use unnumbered third level headings for the acknowledgments. All acknowledgments
% go at the end of the paper. Do not include acknowledgments in the anonymized
% submission, only in the final paper.
